Code Club Project: Assessing whether intra and inter-genomic variation hinder utility of ESVs and ASVs

### 2021-01-20 submission/manuscript.Rmd
Plot number of ASVs per rrn copy for each species by number of genomes

To investigate the variation in the number of copies of the 16S rRNA gene per genome as well as the intragenomic variation among copies of the 16S rRNA gene, I obtained reference 16S rRNA sequences from the rrn copy number database (rrnDB; CITATION). Among the `r n_distinct(metadata$species)` species represented in the rrnDB there were `r nrow(metadata)` genomes. The median number of rrn operson per species ranged between `r single_copy$median` (e.g. *`r single_copy$species`*) and `r max_copy$median` (*`r max_copy$species`*) copies of the rrn operon. As the number of copies of the operon in a genome increased, the number of variants of the 16S rRNA gene in each genome also increased (**FIGURE**). On average, there were `r rate_v19` variants per copy of the full length 16S rRNA gene and an average of `r rate_v4`, `r rate_v34`, and `r rate_v45` variants when considering the V4, V3-V4, and V4-V5 regions of the gene, respectively. Although a species tended to have a consistent number of 16S rRNA gene copies per genome, the number of total variants increased with the number of genomes that were sampled (**FIGURE**). For example, *`r single_copy$species`* generally only had `r single_copy$median` copy of the gene per genome, but across the `r single_copy_N` genomes that have been sequenced there were `r single_nasvs` versions of the gene. Similarly, a *E. coli* genome typically had `r ecoli_copy$median` copies of the 16S rRNA gene with between `r ecoli_copy$min` and `r ecoli_copy$max` distinct full length sequences per genome. Across the `r ecoli_copy$N` *E. coli* genomes that have been sequenced, there were `r ecoli_nasvs` different variants of the gene. These observations highlight the risk of selecting a threshold for defining units of inference that is too narrow because it is possible to split a single genome into multiple units.

@@

## Figures

>>>>

To investigate the variation in the number of copies of the 16S rRNA gene per genome as well as the intragenomic variation among copies of the 16S rRNA gene, I obtained reference 16S rRNA sequences from the rrn copy number database (rrnDB; CITATION). Among the `r n_distinct(metadata$species)` species represented in the rrnDB there were `r nrow(metadata)` genomes. The median number of rrn operson per species ranged between `r single_copy$median` (e.g. *`r single_copy$species`*) and `r max_copy$median` (*`r max_copy$species`*) copies of the rrn operon. As the number of copies of the operon in a genome increased, the number of variants of the 16S rRNA gene in each genome also increased (**FIGURE**). On average, there were `r rate_v19` variants per copy of the full length 16S rRNA gene and an average of `r rate_v4`, `r rate_v34`, and `r rate_v45` variants when considering the V4, V3-V4, and V4-V5 regions of the gene, respectively. Although a species tended to have a consistent number of 16S rRNA gene copies per genome, the number of total variants increased with the number of genomes that were sampled (**Figure 1**). For example, *`r single_copy$species`* generally only had `r single_copy$median` copy of the gene per genome, but across the `r single_copy_N` genomes that have been sequenced there were `r single_nasvs` versions of the gene. Similarly, a *E. coli* genome typically had `r ecoli_copy$median` copies of the 16S rRNA gene with between `r ecoli_copy$min` and `r ecoli_copy$max` distinct full length sequences per genome. Across the `r ecoli_copy$N` *E. coli* genomes that have been sequenced, there were `r ecoli_nasvs` different variants of the gene. These observations highlight the risk of selecting a threshold for defining units of inference that is too narrow because it is possible to split a single genome into multiple units.

@@

\includegraphics{../figures/esv_rate.pdf}

**Figure 1. ESV rate increases as the number of genomes sampled per species increases**

### 2021-01-20 submission/manuscript.tex
Plot number of ASVs per rrn copy for each species by number of genomes

To investigate the variation in the number of copies of the 16S rRNA
gene per genome as well as the intragenomic variation among copies of
the 16S rRNA gene, I obtained reference 16S rRNA sequences from the rrn
copy number database (rrnDB; CITATION). Among the \textbf{4,774} species
represented in the rrnDB there were \textbf{15,614} genomes. The median
number of rrn operson per species ranged between \textbf{1}
(e.g.~\textbf{\emph{Mycobacterium tuberculosis}}) and \textbf{19}
(\textbf{\emph{Metabacillus litoralis}}) copies of the rrn operon. As
the number of copies of the operon in a genome increased, the number of
variants of the 16S rRNA gene in each genome also increased
(\textbf{FIGURE}). On average, there were \textbf{1} variants per copy
of the full length 16S rRNA gene and an average of \textbf{0.26},
\textbf{0.33}, and \textbf{0.27} variants when considering the V4,
V3-V4, and V4-V5 regions of the gene, respectively. Although a species
tended to have a consistent number of 16S rRNA gene copies per genome,
the number of total variants increased with the number of genomes that
were sampled (\textbf{FIGURE}). For example, \textbf{\emph{Mycobacterium
tuberculosis}} generally only had \textbf{1} copy of the gene per
genome, but across the \textbf{180} genomes that have been sequenced
there were \textbf{11} versions of the gene. Similarly, a \emph{E. coli}
genome typically had \textbf{7} copies of the 16S rRNA gene with between
\textbf{6} and \textbf{10} distinct full length sequences per genome.
Across the \textbf{958} \emph{E. coli} genomes that have been sequenced,
there were \textbf{1,013} different variants of the gene. These
observations highlight the risk of selecting a threshold for defining
units of inference that is too narrow because it is possible to split a
single genome into multiple units.

@@

\hypertarget{figures}{%
\subsection{Figures}\label{figures}}

>>>>

To investigate the variation in the number of copies of the 16S rRNA
gene per genome as well as the intragenomic variation among copies of
the 16S rRNA gene, I obtained reference 16S rRNA sequences from the rrn
copy number database (rrnDB; CITATION). Among the \textbf{4,774} species
represented in the rrnDB there were \textbf{15,614} genomes. The median
number of rrn operson per species ranged between \textbf{1}
(e.g.~\textbf{\emph{Mycobacterium tuberculosis}}) and \textbf{19}
(\textbf{\emph{Metabacillus litoralis}}) copies of the rrn operon. As
the number of copies of the operon in a genome increased, the number of
variants of the 16S rRNA gene in each genome also increased
(\textbf{FIGURE}). On average, there were \textbf{1} variants per copy
of the full length 16S rRNA gene and an average of \textbf{0.26},
\textbf{0.33}, and \textbf{0.27} variants when considering the V4,
V3-V4, and V4-V5 regions of the gene, respectively. Although a species
tended to have a consistent number of 16S rRNA gene copies per genome,
the number of total variants increased with the number of genomes that
were sampled (\textbf{Figure 1}). For example,
\textbf{\emph{Mycobacterium tuberculosis}} generally only had \textbf{1}
copy of the gene per genome, but across the \textbf{180} genomes that
have been sequenced there were \textbf{11} versions of the gene.
Similarly, a \emph{E. coli} genome typically had \textbf{7} copies of
the 16S rRNA gene with between \textbf{6} and \textbf{10} distinct full
length sequences per genome. Across the \textbf{958} \emph{E. coli}
genomes that have been sequenced, there were \textbf{1,013} different
variants of the gene. These observations highlight the risk of selecting
a threshold for defining units of inference that is too narrow because
it is possible to split a single genome into multiple units.

@@

\includegraphics{../figures/esv_rate.pdf}

\textbf{Figure 1. ESV rate increases as the number of genomes sampled
per species increases}
